Lossy compression of DNA sequencing quality data

### 2016-04-25 bmc_manuscript/manuscript.tex

\author[
   addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
%   noteref={n1},                        % id's of article notes, if any
   email={chrismh@uw.edu}   % email address
]{\inits{CMH}\fnm{Christopher M} \snm{Hill}}
\author[
   addressref={aff4,aff6},
   email={szolek@informatik.uni-tuebingen.de}
]{\inits{AS}\fnm{Andr\'{a}s} \snm{Szolek}}
\author[
   addressref={aff5,aff6},
   email={mohamed.el-hadidi@uni-tuebingen.de}
]{\inits{MEH}\fnm{Mohamed} \snm{El Hadidi}}
\author[
   addressref={aff3},
   corref={aff2},                       % id of corresponding address, if any
   email={mike@umiac.umd.edu}
]{\inits{MPC}\fnm{Michael P} \snm{Cummings}}

@@

For comparison to other developed methods for lossy compression of
quality values we analyzed the reference data with
\textsc{QualComp}~\cite{Ochoa:2013rt},
\textsc{rsa}~\cite{DBLP:conf/recomb/YuYB14}, and
\textsc{qvz}~\cite{Malysa01102015}.

@@

\textcolor{red}{Add brief description of \textsc{rsa} and
  \textsc{qvz}, including economic costs.}

>>>>

\author[
   addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
%   noteref={n1},                        % id's of article notes, if any
   email={chrismh@uw.edu}   % email address
]{\inits{CMH}\fnm{Christopher M.} \snm{Hill}}
\author[
   addressref={aff4,aff6},
   email={szolek@informatik.uni-tuebingen.de}
]{\inits{AS}\fnm{Andr\'{a}s} \snm{Szolek}}
\author[
   addressref={aff5,aff6},
   email={mohamed.el-hadidi@uni-tuebingen.de}
]{\inits{MEH}\fnm{Mohamed} \snm{El Hadidi}}
\author[
   addressref={aff3},
   corref={aff2},                       % id of corresponding address, if any
   email={mike@umiac.umd.edu}
]{\inits{MPC}\fnm{Michael P.} \snm{Cummings}}

@@

For comparison to other developed methods for lossy compression of
quality values we analyzed the reference data with
\textsc{QualComp}~\cite{Ochoa:2013rt},
\textsc{rqs}~\cite{DBLP:conf/recomb/YuYB14}, and
\textsc{qvz}~\cite{Malysa01102015}.

@@

\textcolor{red}{Add brief description of \textsc{rqs} and
  \textsc{qvz}, including economic costs.}

\textsc{qvz} extends the idea of using rate-distortion theory introduced in \textsc{QualComp}, allowing additional distortion metrics besides \textsc{mse} to be minimized with respect to user-specified rate (bits/read).

\textsc{rqs} first constructs a dictionary of commonly occurring $k$-mers from either the sequencing data set they seek to compress or a larger population-sized sequencing data set.
During compression, $k$-mers within reads that have a small Hamming distance from a dictionary $k$-mer are identified.
These positions may correspond to SNPs or sequencing error.
Positions within the $k$-mer that differ from the nearest $k$-mer in the dictionary are marked.
Lastly, the quality values of the remaining positions are smoothed.
